---
title: "Anomaly Detection"
author: "Duncan Wahira"
date: "7/28/2020"
output: pdf_document
---


## 1. Define the Question


### a. Specifying the question.

As Data Scientists, we have been tasked by Carrefour Kenya with building a model that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).More spefically detect anomalies in sales.

### b. Define the Metrics of Success.

Our metrics of Success is to be able to detect anomalies in sales.

### c. Understand context

Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).

### d. Record the experiment design 

Below is the expected workflow of our Analysis

1.   Data Cleaning and Preparation

*   Load libraries and dataset
*   Deal with missing values
  
2.  Perform EDA

3.  Analysis  

*  Anomaly detection

### e. Data relevance

To establish if the data is relevant to the question and the objectives of the experiment.

Below is the link to the dataset

http://bit.ly/CarreFourSalesDataset

## Load Libraries


```{r}

library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(pander)
library(tidyverse)
library(anomalize)
library(lubridate)
```



## Loading our dataset

```{r}
df <- read.csv("F:\\DATA\\Supermarket_Sales_Forecasting - Sales.csv")
head(df)
```


convert Date column to Date data type

```{r}
df$Date <- as.Date(df$Date,  tryFormats = c("%m-%d-%Y", "%m/%d/%Y"))

str(df$Date)
```


## Data Preprocessing

For Anomaly Detection using anomalize, we need to have either a tibble or tibbletime object. Hence we have to convert the dataframe df into a tibble object that follows a time series shape and store it in df_tidy.

```{r}
df_tidy <- df %>% rownames_to_column() %>% as.tibble() %>% 
  mutate(date = as.Date(Date)) %>% select(-one_of('Date'))
```


looking at the head of df_tidy to see sample data

```{r}
head(df_tidy)
```


```{r}
df_mutate <- df_tidy %>%
  mutate(year = year(date), month = month(date), day = day(date)) %>%
  group_by(year, month, day)%>%
  summarise(sums = mean(Sales))

df_mutate
```



```{r}
plot(sums ~ day, data=df_mutate, type = 'o')

```



## Analysis

Time Series Decomposition with Anomalies


One of the important things to do with Time Series data before starting with Time Series forecasting or Modelling is Time Series Decomposition where the Time series data is decomposed into Seasonal, Trend and remainder components. anomalize has got a function time_decompose() to perform the same. Once the components are decomposed, anomalize can detect and flag anomalies in the decomposed data of the reminder component which then could be visualized with plot_anomaly_decomposition() 



```{r}

df_tidy %>% 
  time_decompose(Sales, method = "stl", frequency = "auto", trend = "auto") %>%
  anomalize(remainder, method = "gesd", alpha = 0.05, max_anoms = 0.2) %>%
  plot_anomaly_decomposition()

```








