---
title: "Dimensionality Reduction"
author: "Duncan Wahira"
date: "7/27/2020"
output: pdf_document
---

## 1. Define the Question


### a. Specifying the question.

As Data Scientists, we have been tasked by Carrefour Kenya with building a model that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).

### b. Define the Metrics of Success.

Our metrics of Success is to be able to do a conclusive Dimensionality Reduction through various methods.

### c. Understand context

Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).

### d. Record the experiment design 

Below is the expected workflow of our Analysis

1.   Data Cleaning and Preparation

*   Load libraries and dataset
*   Deal with missing values
  
2.  Perform EDA

3.  Analysis  

* Seature selection using Filter Methods.

### e. Data relevance

To establish if the data is relevant to the question and the objectives of the experiment.

Below is the link to the dataset

http://bit.ly/SupermarketDatasetII

## Load Libraries

```{r}

library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(pander)
library(Hmisc)
library(tidyverse)
library(corrplot)
library(caret)

```



## Loading our dataset

```{r}
# Preview top 6 rows of data dataset

sales <- read.csv("F:\\DATA\\Supermarket_Dataset_1 - Sales Data.csv")
head(sales)

```

## Checking the data

```{r}
# Preview bottom 6 rows of 

tail(sales)

```

Check shape of our dataset

```{r}

dim(sales)

```
our dataset has 1000 rows and 16 columns


Check column names of Columns Present

```{r}
colnames(sales)
```


shows the data types of the data frame

```{r}

glimpse(sales)

```

Our dataset has 8 numerical variables and 8 factor variable.


Check summary of our dataset

```{r}
summ <- pander(summary(sales))
summ
```
This give of every variable there mean, max,min, datatype and quantiles for numerical variables and Length class and Mode for factor variables.


## External Data Source Validation

Making sure your data matches something outside of the dataset is very important. It allows you to ensure that the measurements are roughly in line with what they should be and it serves as a check on what other things might be wrong in your dataset. External validation can often be as simple as checking your data against a single number, as we will do here.


## Tidying the Dataset

We start by checking for null values in our dataset

```{r}

# Check for missing values

colSums(is.na(sales))

```
Our dataset has no null values


Remove Duplicated values
```{r}
dt <- unique(sales)
dim(dt)
```
There are no duplicated values


Drop ID column

```{r}
dt$Invoice.ID <- NULL

```



convert Factors variables to Numeric 

```{r}

dt_numeric = dt

dt_numeric$Date <- NULL
dt_numeric$Time <- NULL
dt_numeric$gross.margin.percentage <- NULL


conv <- c("Branch", "Customer.type", "Gender", "Product.line", "Payment")

dt_numeric$Branch <- as.numeric(as.factor(dt_numeric$Branch))
dt_numeric$Customer.type <- as.numeric(as.factor(dt_numeric$Customer.type))
dt_numeric$Gender <- as.numeric(as.factor(dt_numeric$Gender))
dt_numeric$Product.line <- as.numeric(as.factor(dt_numeric$Product.line))
dt_numeric$Payment <- as.numeric(as.factor(dt_numeric$Payment))
glimpse(dt_numeric)
```
convert factors variables to numeric

convert Date variables to Date
```{r}

dt$Date <- as.Date(dt$Date,  tryFormats = c("%m-%d-%Y", "%m/%d/%Y"))

str(dt$Date)

```


## Feature Selection

Method 1 - Filter Methods  


```{r}
correlationMatrix <- cor(dt_numeric)


```


Find attributes that are highly correlated

```{r}

highlyCorrelated <- findCorrelation(correlationMatrix, cutoff=0.90)
names(dt_numeric[,highlyCorrelated])
```


We can remove the variables with a higher correlation and comparing the results graphically as shown .

```{r}

dt_clean <- dt_numeric[-highlyCorrelated]

```

Performing our graphical comparison

```{r}

par(mfrow = c(1, 2))
corrplot(correlationMatrix, order = "hclust")
corrplot(cor(dt_clean), order = "hclust")

```

View Correlation Matrix

```{r}

correlation <- data.frame(correlationMatrix)
correlation

```


Extract product line correlations

```{r}
product <- data.frame(correlation$Product.line, row.names = rownames(correlation))
product

```

Select Features

```{r}

Featureselected <- data.frame(product[product["correlation.Product.line"] > 0],row.names = c("Branch","Gender","Product.line","Unit.price", "Quantity", "Tax", "Payment", "cogs", "gross.income"))
Featureselected

```

